\documentclass[10pt]{article}
\usepackage[T1]{fontenc}

% Document Details
\newcommand{\CLASS}{AMATH 514}
\newcommand{\assigmentnum}{Assignment 9}

\usepackage[margin = 1.15in, top = 1.25in, bottom = 1.in]{geometry}

\input{../../TeX_headers/title.tex} % Title Styling
\input{../../TeX_headers/sfftoc.tex} % General Styling
\input{../../TeX_headers/styling.tex} % General Styling
\input{../../TeX_headers/code.tex} % Code Display Setup
\input{../../TeX_headers/math.tex} % Math shortcuts
\input{../../TeX_headers/problem.tex} % Problem Environment

\newcommand{\note}[1]{\textcolor{red}{\textbf{Note:} #1}}

\hypersetup{
   colorlinks=true,       % false: boxed links; true: colored links
   linkcolor=violet,          % color of internal links (change box color with linkbordercolor)
   citecolor=green,        % color of links to bibliography
   filecolor=magenta,      % color of file links
   urlcolor=cyan           % color of external links
}


\begin{document}
\maketitle



\begin{problem}[Problem 10.5]
Let \( M = (X, \mathcal{I} ) \) be a matroid and let \( k \) be a natural number. Let \( \mathcal{I}'  := \{Y\subseteq \mathcal{I}  : |Y| \leq k\} \). Show that \( (X, \mathcal{I}' ) \) is again a matroid.
\end{problem}

\begin{solution}

\begin{enumerate}[label=(\roman*)]
    \item Since \( (X, \mathcal{I} ) \) is a matroid then \( \emptyset \in \mathcal{I} \). Clearly \( |\emptyset| \leq k \) so \( \emptyset \in \mathcal{I} ' \).
    \item Let \( Y \in \mathcal{I}' \) and \( Z\subseteq Y \). Since \( (X, \mathcal{I} ) \) is a matroid then \( Z\in \mathcal{I} \).  Moreover, \( |Z| \leq |Y| \leq k \) as \( Z\in \mathcal{I} ' \). Therefore \( Z\in \mathcal{I} ' \).
    \item Let \( Y,Z\in  \mathcal{I}' \) with \( |Y|<|Z| \). Since \( (X, \mathcal{I} ) \) is a matroid and \( Y,Z\in \mathcal{I} \) then \( Y\cup \{x\} \in \mathcal{I} \) for some \( x\in Z \setminus Y \). Since \( Z\in \mathcal{I} ' \), \( |Z|\leq k \). Therefore \( |Y\cup \{x\}| = |Y|+ 1 \leq |Z| \leq k \). Therefore \( Y\cup\{x\} \in \mathcal{I} ' \).
\end{enumerate}

This proves \( (X, \mathcal{I} ') \) is a matroid. \qed

\end{solution}

\begin{problem}[Problem 10.19]
    Let \( M = (X, \mathcal{I} ) \) be a matroid, let \( B \) be a basis of \( M \), and let \( w: X\to\RR \) be a weight function. Show that \( B \) is a basis of maximum weight if and only if \( w(B') \leq w(B) \) for every basis \( B' \) with \( |B' \setminus B| = 1 \).
\end{problem}

\begin{solution}

Let \( B,B' \) be bases. From theorem we have:
\begin{enumerate}[label=(\roman*),nolistsep]
    \item for any \( x\in B'\setminus B \), \( (B'\setminus\{x\})\cup \{y\} \) is a basis of \( M \) for some  \( y\in B\setminus B' \).
    \item for any \( x\in B'\setminus B \), \( (B\setminus\{y\})\cup \{x\} \) is a basis of \( M \) for some  \( y\in B\setminus B' \).
\end{enumerate}


\vspace{1em}
Suppose \( B \) is a basis of maximum weight and let \( B' \) be some basis (with \( |B'\setminus B| = 1 \)). Then \( w(B') \leq w(B) \). 

Conversely, suppose that \( w(B^\dagger) \leq w(B) \) for every basis \( B^\dagger \) with \( |B^\dagger \setminus B| = 1 \). 

Let \( B' \) be a basis of \( M \).

Suppose \( | B' \setminus B| = 0 \). Then, since all bases have the same size we have \( B' = B \) so \( w(B') \leq w(B) \).

Now, suppose \( |B' \setminus B| > 0 \).

We induct on \( k := |B' \setminus B| \), assuming that \( w(B^\dagger) \leq w(B) \) for all bases \( B^\dagger \) of \( M \) with \( |B^\dagger \setminus B| < k \). Clearly the original hypothesis is the base case of our induction.

Since \( |B'\setminus B| > 0 \) there is some \( x\in B'\setminus B \). Therefore, by the theorem listed above there is some \( y\in B\setminus B' \) such that \( B^\dagger = (B' \setminus \{x\}) \cup \{y\} \) is a basis for \( M \).

Observe,
\begin{align*}
    |B^\dagger\setminus B| = | ((B'\setminus \{ x \})\cup\{y\}) \setminus B| = | (B'\setminus B)\setminus\{x\} | = k - 1 
\end{align*}

Therefore, by our induction hypothesis we have,
\begin{align*}
    w(B') - w(x) + w(y) = w(B^\dagger) \leq w(B)
\end{align*}


Since,
\begin{align*}
    |((B\setminus \{y\})\cup\{x\}) \setminus B| = |\{x\}| = 1 
\end{align*}
by the base case we have,
\begin{align*}
    w((B\setminus\{y\})\cup\{x\}) \leq w(B)
\end{align*}

Therefore, 
\begin{align*}
    w(B') \leq w(B) -w(y) + w(x) = w( (B\setminus\{y\} )\cup \{x\}) \leq w(B)
\end{align*}


This proves the result. \qed

\iffalse

Recall \( |B''| = |B| \) for any two bases \( B'' \) and \( B \) of \( M \). Therefore \( |B''\setminus B| = 0 \) means \( B = B'' \) in which case \( w(B'') = w(B) \).

Start with any basis \( B_0 \) of \( B \) with \( |B_0\setminus B| > 1 \).

At step \( k \in\{0,1,\ldots \} \), if \( |B_k \setminus B| = 0 \) stop.

Then at each step there is some element \( x\in B_k\setminus B \). By theorem we have, \( B_{k+1}:= (B\setminus\{y_k\})\cup\{x_k\} \) a basis for \( M \) for some \( y_k\in B\setminus B_k \).

Note \( |((B\setminus \{y_k\})\cup\{k_0\})\setminus B| = |\{x_k\}| = 1 \). Then, by hypothesis,
\begin{align*}
    w(B) \geq w((B\setminus \{y_k\}) \cup\{x_k\}) = w(B) - w(y_k) + w(x_k)
\end{align*}
Therefore,
\begin{align*}
    w(y_k) \geq w(x_k)
\end{align*}


At the end of the process we will have sets \( X:= \{x_k\}_{k\geq 0} \) and \( Y:= \{y_k\}_{k \geq 0} \) satisfying,
\begin{align*}
    w(Y) \geq w(X)
\end{align*}

We will also have,
\begin{align*}
    B_k = ( B \setminus Y )\cup X = B'
\end{align*}

Therefore,
\begin{align*}
    w(B') = w(B_k) = w(B) - w(Y) + w(X) \leq w(B)
\end{align*}

\fi




\iffalse
Now note, \( |((B'\setminus\{x\})\cup \{y\})\setminus B| = | (B'\setminus B)\setminus\{x\} | = k  \). Therefore, by our induction hypothesis,
\begin{align*}
    w(B) \geq w((B'\setminus\{x\})\cup\{y\}) = 
\end{align*}
\fi


\end{solution}

\begin{problem}[Problem 10.22]
Derive K\"onig's matching theorem from Edmonds' matroid intersection theorem. 
\end{problem}

\begin{solution}

Let \( G = (V_1\cup V_2, E) \) be a bipartite graph.
Let \( \mathcal{I}_1 \) be the collection of all subsets \( F \) of \( E \) so that no two edges in \( F \) have a vertex in \( V_1 \) in common.
Similarly, let \( \mathcal{I}_w \) be the collection of all subsets \( F \) of \( E \) so that no two edges in \( F \) have a vertex in \( V_2 \) in common.
Then both \( M_1:=(E, \mathcal{I}_1)  \) and \( M_2:=(E, \mathcal{I}_2)  \) are partition matroids.

Now observe that elements of \( \mathcal{I}_1\cap \mathcal{I}_2 \) are matchings in \( G \), and matchings in \( G \) are elements of \( \mathcal{I}_1\cap \mathcal{I}_2 \).

Therefore,
\begin{align*}
    \nu(G) = \max_{Y\in \mathcal{I}_1\cap \mathcal{I}_2} |Y| 
\end{align*}


Let \( U\subseteq E \). Define,
\begin{align*}
    C_1 = \{ v\in V_1 : \exists e\in E \text{ with } v\in e \}
\end{align*}

Now define,
\begin{align*}
    C_2 = \{ v\in V_2 : v \text{ reachable from } V_1\setminus C_1 \}
\end{align*}

Finally, let \( C = C_1 \cup C_2 \). If \( e\in U \) then \( e\cap C_1 \neq \emptyset \). Similarly, if \( e\in E\setminus U \) then \( e\cap C_2 \neq \emptyset \). Therefore \( C \) is a vertex cover of \( G \).

Every edge in \( U \) touches a vertex in \( C_1 \), and every vertex in \( C_1 \) touches an edge in \( U \). 
Similarly, every edge in \( E\setminus U \) touches a vertex in \( C_2 \), and every vertex in \( C_2 \) touches an edge in \( E\setminus U \). 

Therefore \( r_{M_1}(U) = |C_1| \) and \( r_{M_2}(E\setminus U) = |C_2| \) so that \( r_{M_1}(U) + r_{M_2}(E\setminus U) = |C| \).

In particular, this proves,
\begin{align*}
    \min_{U\subseteq E} (r_{M_1}(U) + r_{M_2}(E\setminus U)) \geq  \tau(G)
\end{align*}




Now, let \( C \) be a minimum size vertex cover of \( G \). Define,
\begin{align*}
    U = \{ e\in E: (C\cap V_1)\cap e \neq \emptyset  \}
\end{align*}

Therefore every edge in \( U \) is covered by some vertex in \( C\cap V_1 \) and every vertex in \( C\cap V_1 \) touches some edge in \( U \). This means that that \( r_{M_1}(U) = |C\cap V_1| \).

Similarly, every edge in \( E\setminus U \) must be covered by some vertex in \( C \). Moreover, by the minimality of \( C \), every vertex in \( C \) must touch an edge in \( E\setminus U \) (otherwise \( C \) without this edge would be a smaller vertex cover). Therefore \( r_{M_2}(E\setminus U) = |C\cap V_2| \)

Since points in \( C \) are either in \( V_1 \) or \( V_2 \) we have found a \( U\subseteq E \) such that,
\begin{align*}
    r_{M_1}(U) + r_{M_2}(E\setminus U) = |C| = \tau(G)
\end{align*}

This proves,
\begin{align*}
    \min_{U\subseteq E} (r_{M_1}(U) + r_{M_2}(E\setminus U)) \leq  \tau(G)
\end{align*}



Using Edmond's matroid intersection theorem we have,
\begin{align*}
    \nu(G) 
    = \max_{Y\in \mathcal{I}_1\cap \mathcal{I}_2} |Y| 
    = \min_{U\subseteq X} (r_{M_1}(U) + r_{M_2}(E\setminus U)) 
    = \tau(G)
\end{align*}

This is Konig's matching theorem. \qed

\end{solution}

\end{document}
